---
title: "Analysis of ENCEVI 2018"
author: '[Mauricio Hernandez](http://mauricioh2.com)'
date: "`r Sys.Date()`"
subtitle: Getting Consumption Statistics
bibliography: energy_mex.bib
---




```{r, include=FALSE}
#Remove all objects from current workspace and call garbage collector
rm(list=ls())
gc()

source("./script/general_setup_functions.R")
source("./script/compute_consumption_bill.R")
source("./script/get_bill_periodicity.R")
```

Notes: Dollar Exchange Rate: 	18.70	15-Mar-18
```{r}
DOLLAR_TO_MXN <- 18.70
```

```{r knitr_init, include=FALSE}
#options(max.print = "75")
opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               comment = NA,
               message = FALSE,
               warning = FALSE,
               fig.path = 'figures/')
```

```{r setup, include=FALSE}
#Generates the css needed by summarytools
st_css()

#devtools::install_github("ropensci/plotly")
knitr::opts_chunk$set(fig.width = 8, collapse = TRUE)

knitr::knit_hooks$set(output = function(x, options){
  if(!is.null(options$max_height)){
    paste('<pre style = "max-height:',
          options$max_height, 
          ';float: left; width: 910px; overflow-y: auto;">',
          x,
          "</pre>",
          sep = "")
  } else{
    x
  }
})
```

**Loading datasets**
</br>This dataset was obtained with the "05_bill_power_calculation.Rmd" file.
```{r}
df.bill.cons <- read.csv("./output/df_bill_wconsump.csv", 
                           header = TRUE, na.strings="")
```


##Calculating statistics of data that didn't have to be recovered
**Monthly consumption by region and summer period, data that hasn't been recovered**
```{r, results="hold"}
df.bill.cons.temp <- subset(df.bill.cons, df.bill.cons$day.recovered == 0) %>%
#  group_by(bill.period1, mes_inic1.f) %>% 
  group_by(is.summer1, region.f) %>% 
  summarize(Min = min(cons.month.kWh.med1, na.rm=TRUE),
    q1 = quantile(cons.month.kWh.med1, 0.25, na.rm=TRUE),
    Median=median(cons.month.kWh.med1, na.rm=TRUE),
    Mean=mean(cons.month.kWh.med1, na.rm=TRUE),
    q3 = quantile(cons.month.kWh.med1, 0.75, na.rm=TRUE),
    Max = max(cons.month.kWh.med1, na.rm=TRUE),
    sd=sd(bill.amount1, na.rm=TRUE),
    n = sum(!is.na(cons.month.kWh.med1)),
    NAs = sum(is.na(cons.month.kWh.med1))
    )

df.bill.cons.temp

write.csv(df.bill.cons.temp, "./output/TEMP_bill_cons_summer_NOrecov.csv", 
          row.names = FALSE, na="")
```
**Monthly electricity consumption by region and summer period, data that was recovered**
```{r, results="hold"}
df.bill.cons.temp <- subset(df.bill.cons, df.bill.cons$day.recovered == 1) %>%
#  group_by(bill.period1, mes_inic1.f) %>% 
  group_by(is.summer1, region.f) %>% 
  summarize(Min = min(cons.month.kWh.med1, na.rm=TRUE),
    q1 = quantile(cons.month.kWh.med1, 0.25, na.rm=TRUE),
    Median=median(cons.month.kWh.med1, na.rm=TRUE),
    Mean=mean(cons.month.kWh.med1, na.rm=TRUE),
    q3 = quantile(cons.month.kWh.med1, 0.75, na.rm=TRUE),
    Max = max(cons.month.kWh.med1, na.rm=TRUE),
    sd=sd(bill.amount1, na.rm=TRUE),
    n = sum(!is.na(cons.month.kWh.med1)),
    NAs = sum(is.na(cons.month.kWh.med1))
    )

df.bill.cons.temp

write.csv(df.bill.cons.temp, "./output/TEMP_bill_cons_summer_recov.csv", 
          row.names = FALSE, na="")  
```

**Monthly electricity consumption by region and summer period, for all the valid data (data that was recovered and that didn't have to be recovered)**
```{r, results="hold"}
df.bill.cons.temp <- subset(df.bill.cons) %>%
#  group_by(bill.period1, mes_inic1.f) %>% 
  group_by(is.summer1, region.f) %>% 
  summarize(Min = min(cons.month.kWh.med1, na.rm=TRUE),
    q1 = quantile(cons.month.kWh.med1, 0.25, na.rm=TRUE),
    Median=median(cons.month.kWh.med1, na.rm=TRUE),
    Mean=mean(cons.month.kWh.med1, na.rm=TRUE),
    q3 = quantile(cons.month.kWh.med1, 0.75, na.rm=TRUE),
    Max = max(cons.month.kWh.med1, na.rm=TRUE),
    sd=sd(bill.amount1, na.rm=TRUE),
    n = sum(!is.na(cons.month.kWh.med1)),
    NAs = sum(is.na(cons.month.kWh.med1))
    )

df.bill.cons.temp

write.csv(df.bill.cons.temp, "./output/TEMP_bill_cons_summer_all.csv", 
          row.names = FALSE, na="")  
```

## Payments per KWh after taxes

Price per Kwh 
```{r}
df.bill.cons$price.kWh.med1 <- df.bill.cons$bill.amount1 / df.bill.cons$cons.kWh.med1
```

**Amount of MXN$ paid by KWh by region, data that didn't have to be recovered**
```{r, results="hold"}
df.bill.cons.temp <- subset(df.bill.cons, df.bill.cons$day.recovered == 0) %>%
  group_by(is.summer1, region.f) %>% 
  summarize(Min = min(price.kWh.med1, na.rm=TRUE),
    q1 = quantile(price.kWh.med1, 0.25, na.rm=TRUE),
    Median=median(price.kWh.med1, na.rm=TRUE),
    Mean=mean(price.kWh.med1, na.rm=TRUE),
    q3 = quantile(price.kWh.med1, 0.75, na.rm=TRUE),
    Max = max(price.kWh.med1, na.rm=TRUE),
    sd=sd(price.kWh.med1, na.rm=TRUE),
    n = sum(!is.na(price.kWh.med1)),
    NAs = sum(is.na(price.kWh.med1))
    )

df.bill.cons.temp

write.csv(df.bill.cons.temp, "./output/SUM_Bill_pricekwh_summer_NOrecov.csv", 
          row.names = FALSE, na="")
```


**Amount of MXN$ paid by KWh by region, data that was recovered**
```{r, results="hold"}
df.bill.cons.temp <- subset(df.bill.cons, df.bill.cons$day.recovered == 1) %>%
  group_by(is.summer1, region.f) %>% 
  summarize(Min = min(price.kWh.med1, na.rm=TRUE),
    q1 = quantile(price.kWh.med1, 0.25, na.rm=TRUE),
    Median=median(price.kWh.med1, na.rm=TRUE),
    Mean=mean(price.kWh.med1, na.rm=TRUE),
    q3 = quantile(price.kWh.med1, 0.75, na.rm=TRUE),
    Max = max(price.kWh.med1, na.rm=TRUE),
    sd=sd(price.kWh.med1, na.rm=TRUE),
    n = sum(!is.na(price.kWh.med1)),
    NAs = sum(is.na(price.kWh.med1))
    )

df.bill.cons.temp

write.csv(df.bill.cons.temp, "./output/SUM_bill_pricekwh_summer_recov.csv", 
          row.names = FALSE, na="")
```

**Amount of MXN$ paid by KWh by region, data that was recovered**
```{r, results="hold"}
df.bill.cons.temp <- subset(df.bill.cons) %>%
  group_by(is.summer1, region.f) %>% 
  summarize(Min = min(price.kWh.med1, na.rm=TRUE),
    q1 = quantile(price.kWh.med1, 0.25, na.rm=TRUE),
    Median=median(price.kWh.med1, na.rm=TRUE),
    Mean=mean(price.kWh.med1, na.rm=TRUE),
    q3 = quantile(price.kWh.med1, 0.75, na.rm=TRUE),
    Max = max(price.kWh.med1, na.rm=TRUE),
    sd=sd(price.kWh.med1, na.rm=TRUE),  
    n = sum(!is.na(price.kWh.med1)),
    NAs = sum(is.na(price.kWh.med1))
    )

df.bill.cons.temp

write.csv(df.bill.cons.temp, "./output/SUM_bill_pricekwh_summer_all.csv",
          row.names = FALSE, na="")
```    
          

## Payments per KWh before taxes

```{r}
df.bill.cons$price.kWh.notax.med1 <- df.bill.cons$bill.avg1.notax / df.bill.cons$cons.kWh.med1
```


**Amount of MXN$ paid by KWh by region (before taxes), data that didn't have to be recovered**
```{r, results="hold"}
df.bill.cons.temp <- subset(df.bill.cons, df.bill.cons$day.recovered == 0) %>%
  group_by(is.summer1, region.f) %>% 
  summarize(Min = min(price.kWh.notax.med1, na.rm=TRUE),
    q1 = quantile(price.kWh.notax.med1, 0.25, na.rm=TRUE),
    Median=median(price.kWh.notax.med1, na.rm=TRUE),
    Mean=mean(price.kWh.notax.med1, na.rm=TRUE),
    q3 = quantile(price.kWh.notax.med1, 0.75, na.rm=TRUE),
    Max = max(price.kWh.notax.med1, na.rm=TRUE),
    sd=sd(price.kWh.notax.med1, na.rm=TRUE),
    n = sum(!is.na(price.kWh.notax.med1)),
    NAs = sum(is.na(price.kWh.notax.med1))
    )

df.bill.cons.temp

write.csv(df.bill.cons.temp, "./output/SUM_Bill_pricekwhNOTAX_summer_NOrecov.csv", 
          row.names = FALSE, na="")
```


**Amount of MXN$ paid by KWh by region, data that was recovered**
```{r, results="hold"}
df.bill.cons.temp <- subset(df.bill.cons, df.bill.cons$day.recovered == 1) %>%
  group_by(is.summer1, region.f) %>% 
  summarize(Min = min(price.kWh.notax.med1, na.rm=TRUE),
    q1 = quantile(price.kWh.notax.med1, 0.25, na.rm=TRUE),
    Median=median(price.kWh.notax.med1, na.rm=TRUE),
    Mean=mean(price.kWh.notax.med1, na.rm=TRUE),
    q3 = quantile(price.kWh.notax.med1, 0.75, na.rm=TRUE),
    Max = max(price.kWh.notax.med1, na.rm=TRUE),
    sd=sd(price.kWh.notax.med1, na.rm=TRUE),
    n = sum(!is.na(price.kWh.notax.med1)),
    NAs = sum(is.na(price.kWh.notax.med1))
    )

df.bill.cons.temp

write.csv(df.bill.cons.temp, "./output/SUM_bill_pricekwhNOTAX_summer_recov.csv", 
          row.names = FALSE, na="")
```

**Amount of MXN$ paid by KWh by region, data that was recovered**
```{r, results="hold"}
df.bill.cons.temp <- subset(df.bill.cons) %>%
  group_by(is.summer1, region.f) %>% 
  summarize(Min = min(price.kWh.notax.med1, na.rm=TRUE),
    q1 = quantile(price.kWh.notax.med1, 0.25, na.rm=TRUE),
    Median=median(price.kWh.notax.med1, na.rm=TRUE),
    Mean=mean(price.kWh.notax.med1, na.rm=TRUE),
    q3 = quantile(price.kWh.notax.med1, 0.75, na.rm=TRUE),
    Max = max(price.kWh.notax.med1, na.rm=TRUE),
    sd=sd(price.kWh.notax.med1, na.rm=TRUE),  
    n = sum(!is.na(price.kWh.notax.med1)),
    NAs = sum(is.na(price.kWh.notax.med1))
    )

df.bill.cons.temp

write.csv(df.bill.cons.temp, "./output/SUM_bill_pricekwhNOTAX_summer_all.csv",
          row.names = FALSE, na="")
```    

         

```{r, results="hold"}
df.bill.cons.temp <- subset(df.bill.cons, df.bill.cons$day.recovered == 1) %>%
  group_by(is.summer1, region.f) %>% 
  summarize(Min = min(price.kWh.notax.med1, na.rm=TRUE),
    q1 = quantile(price.kWh.notax.med1, 0.25, na.rm=TRUE),
    Median=median(price.kWh.notax.med1, na.rm=TRUE),
    Mean=mean(price.kWh.notax.med1, na.rm=TRUE),
    q3 = quantile(price.kWh.notax.med1, 0.75, na.rm=TRUE),
    Max = max(price.kWh.notax.med1, na.rm=TRUE),
    n = sum(!is.na(price.kWh.notax.med1)),
    NAs = sum(is.na(price.kWh.notax.med1))
    )

df.bill.cons.temp

#write.csv(df.bill.cons.temp, "./output/TEMP_bill_cons_recov.csv", 
#          row.names = FALSE, na="")
```

**Amount of MXN$ paid by KWh by region, data that was recovered**
```{r, results="hold"}
df.bill.cons.temp <- subset(df.bill.cons, df.bill.cons$day.recovered == 1) %>%
  group_by(is.summer1, region.f) %>% 
  summarize(Min = min(bill.amount1, na.rm=TRUE),
    q1 = quantile(bill.amount1, 0.25, na.rm=TRUE),
    Median=median(bill.amount1, na.rm=TRUE),
    Mean=mean(bill.amount1, na.rm=TRUE),
    q3 = quantile(bill.amount1, 0.75, na.rm=TRUE),
    Max = max(bill.amount1, na.rm=TRUE),
    n = sum(!is.na(bill.amount1)),
    NAs = sum(is.na(bill.amount1))
    )

df.bill.cons.temp

#write.csv(df.bill.cons.temp, "./output/TEMP_bill_cons_recov.csv", 
#          row.names = FALSE, na="")
```


**Monthly consumption by region, all data**
```{r, results="hold"}

df.bill.cons.temp <- subset(df.bill.cons) %>%
  group_by(mes_inic1.f) %>% 
  summarize(Min = min(cons.month.kWh.med1, na.rm=TRUE),
    q1 = quantile(cons.month.kWh.med1, 0.25, na.rm=TRUE),
    Median=median(cons.month.kWh.med1, na.rm=TRUE),
    Mean=mean(cons.month.kWh.med1, na.rm=TRUE),
    q3 = quantile(cons.month.kWh.med1, 0.75, na.rm=TRUE),
    Max = max(cons.month.kWh.med1, na.rm=TRUE),
    n = sum(!is.na(cons.month.kWh.med1)),
    NAs = sum(is.na(cons.month.kWh.med1))
    )

df.bill.cons.temp

write.csv(df.bill.cons.temp, "./output/TEMP_bill_cons_all.csv", 
          row.names = FALSE, na="")
```


# TEMP

**Monthly consumption by region, data that hasn't been recovered**


**Bill Amount**

**Owed bill by region, data that was recovered**


```{r}
# Summary by group using dplyr
library(psych)
#psych::describeBy(df.bill.cons.temp$cons.month.kWh.med1, group=df.bill.cons.temp$region, mat=FALSE,type=3, digits=15)

describe.by(df.bill.cons.amount, df.bill.cons.amount$region)
```



```{r}
df.bill.cons.temp <- subset(df.bill.cons, df.bill.cons$day.recovered == 0,
                        select=c('cons.month.kWh.low1', 'cons.month.kWh.med1', 
                                 'cons.month.kWh.hig1', 'region'))
```


```{r}
descr(df.bill.cons$day.recovered)
#df.bill.cons$cons.month.kWh.med1
df.bill.temp <- subset(df.bill.cons,
                        select=c('cons.day.kWh.low1', 'cons.day.kWh.med1', 
                                 'cons.day.kWh.hig1', 'factor_sem'))


#df.bill.cons$final.month.mod[df.bill.cons$day.recovered == 0]

dstrata <- df.bill.cons %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl.mean.cons <- dstrata %>% 
  group_by(SES, day.recovered)  %>%
  summarize(mean_cons_month = survey_mean(cons.month.kWh.med1, na.rm= TRUE, vartype = c("ci")))

```


