---
title: "Analysis of ENCEVI 2018"
author: '[Mauricio Hernandez](http://mauricioh2.com)'
date: "`r Sys.Date()`"
subtitle: Estimate Subsidies
bibliography: energy_mex.bib
---


```{r, include=FALSE}
#Remove all objects from current workspace and call garbage collector
rm(list=ls())
gc()

source("./script/general_setup_functions.R")
source("./script/compute_consumption_bill.R")
source("./script/get_bill_periodicity.R")
```


Notes: Dollar Exchange Rate: 	18.70	30-Jun-18, Ref. Banxico, 2021 (https://www.banxico.org.mx/tipcamb/main.do?page=tip&idioma=sp)
```{r}
DOLLAR_TO_MXN <- 19.69
```

```{r knitr_init, include=FALSE}
#options(max.print = "75")
opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               comment = NA,
               message = FALSE,
               warning = FALSE,
               fig.path = 'figures/')
```

```{r setup, include=FALSE}
#Generates the css needed by summarytools
st_css()

#devtools::install_github("ropensci/plotly")
knitr::opts_chunk$set(fig.width = 8, collapse = TRUE)

knitr::knit_hooks$set(output = function(x, options){
  if(!is.null(options$max_height)){
    paste('<pre style = "max-height:',
          options$max_height, 
          ';float: left; width: 910px; overflow-y: auto;">',
          x,
          "</pre>",
          sep = "")
  } else{
    x
  }
})
```

**Loading datasets**
</br>This dataset was obtained with the "06_getting_consumption_statistics.Rmd" notebook.
```{r}

df.bill.cons <- read.csv("./output/df_bill_wconsump.csv", 
                           header = TRUE, na.strings="")

df.tariff.scheme  <- read.csv("./output/tables/tariff_cost.csv", 
                              header = TRUE, na.strings="")
head(df.tariff.scheme)
```

Actual cost of electricity is similar for each tariff and for summer and no summer seasons
The cost per Kwh is $2.802 MXN
```{r}
df.tariff.scheme.sub <- subset(df.tariff.scheme, tariff.block.name=="extra",
                               select= c(tariff, summer, cost.kwh))
df.tariff.scheme.sub
```

```{r}
# for all the tariffs the max. cost per kwh is $2.802
COST_KWH_CFE <- mean(df.tariff.scheme.sub$cost.kwh)
COST_KWH_CFE_USD <- COST_KWH_CFE / DOLLAR_TO_MXN
```


## Payments per KWh before taxes
```{r}
df.bill.cons$bill.avg1.notax[df.bill.cons$bill.avg1.notax <= min(df.tariff.scheme$cost.kwh)*25] <- NA  

#Price of electricity with subsidies and without subsidies (real price)
#These values don't include taxes and fees
df.bill.cons$price.kwh.notax <- df.bill.cons$bill.avg1.notax / df.bill.cons$cons.kWh.med1
df.bill.cons$price.kwh.notax.usd <- df.bill.cons$price.kwh.notax / DOLLAR_TO_MXN
df.bill.cons$real.price.kwh <- COST_KWH_CFE
df.bill.cons$real.price.kwh.usd <-  COST_KWH_CFE_USD

#Daily cost of electricity

#These values don't include taxes and fees
df.bill.cons$cost.day.notax <- df.bill.cons$price.kwh.notax * df.bill.cons$cons.day.kWh.med1
df.bill.cons$cost.day.notax.usd <- df.bill.cons$cost.day.notax  / DOLLAR_TO_MXN
df.bill.cons$real.cost.day <- df.bill.cons$real.price.kwh * df.bill.cons$cons.day.kWh.med1
df.bill.cons$real.cost.day.usd <- df.bill.cons$real.cost.day / DOLLAR_TO_MXN

#Annual cost of electricity
#These values don't include taxes and fees
df.bill.cons$cost.year.notax < - df.bill.cons$cost.day.notax * 365
df.bill.cons$cost.year.notax.usd <- df.bill.cons$cost.day.notax.usd * 365
df.bill.cons$real.cost.year <- df.bill.cons$real.cost.day * 365
df.bill.cons$real.cost.year.usd <- df.bill.cons$real.cost.day.usd * 365
```


##Average annual consumption
```{r}
options( survey.lonely.psu = "adjust" )
dstrata <- df.bill.cons %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl.month.consumption <- dstrata %>%
  summarize(mean_month_consumption = survey_mean(cons.month.kWh.med1, na.rm= TRUE, vartype = c("ci"))) 

tbl.month.consumption
```

```{r}
40*4
```


```{r}
# just to check calculations
#price weighted 
df.bill.cons$real.cost.year.usd.w <- df.bill.cons$real.cost.year.usd * df.bill.cons$factor_sem
```

##Price per KWh
**Price without subsidies**
```{r}
options( survey.lonely.psu = "adjust" )
dstrata <- df.bill.cons %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl.kwh.price.nosub.usd <- dstrata %>%
  summarize(mean_price_kwh_usd = survey_mean(real.price.kwh.usd, na.rm= TRUE, vartype = c("ci"))) 

tbl.kwh.price.nosub.usd
```

**Price with subsidies**
```{r}
options(survey.lonely.psu = "adjust" )
dstrata <- df.bill.cons %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl.kwh.price.usd <- dstrata %>%
  summarize(mean_price_kwh_usd = survey_mean(price.kwh.notax.usd, na.rm= TRUE, vartype = c("ci"))) 

tbl.kwh.price.usd
```
```{r}
df.bill.cons.sub <- subset(df.bill.cons, select=c(upm, est_dis, factor_sem, price.kwh.notax, price.kwh.notax.usd))

dstrat.sub <- df.bill.cons.sub %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl.kwh.price.sub.usd <- dstrat.sub %>%
  summarize(mean_price_kwh_usd = survey_mean(price.kwh.notax.usd, na.rm= TRUE, vartype = c("ci"))) 

tbl.kwh.price.sub.usd
```

```{r}
descr(df.bill.cons.sub)
```


##Price per KWh by SES
Price with subsidies
```{r}
tbl.price.kwh.ses.usd <- dstrata %>% 
  group_by(SES)  %>%
  summarize(mean_price_kwh_usd = survey_mean(price.kwh.notax.usd, na.rm= TRUE, vartype = c("ci"))) 

tbl.price.kwh.ses.usd <- na.omit(tbl.price.kwh.ses.usd) 

tbl.price.kwh.ses.usd$SES <- as.character(tbl.price.kwh.ses.usd$SES)

tbl.price.kwh.ses.usd
```


##Price per KWh by Region
### Price of KWh per Region without subsidies
```{r}
tbl.price.kwh.region.nosub.usd <- dstrata %>% 
  group_by(region.f)  %>%
  summarize(mean_price_kwh_usd = survey_mean(real.price.kwh.usd, na.rm= TRUE, vartype = c("ci"))) 

tbl.price.kwh.region.nosub.usd <- na.omit(tbl.price.kwh.region.nosub.usd) 

tbl.price.kwh.region.nosub.usd
```

### Price of KWh per Region with subsidies
```{r}
#df.bill.cons$
tbl.price.kwh.region.usd <- dstrata %>% 
  group_by(region.f)  %>%
  summarize(mean_price_kwh_usd = survey_mean(price.kwh.notax.usd, na.rm= TRUE, vartype = c("ci"))) 

tbl.price.kwh.region.usd <- na.omit(tbl.price.kwh.region.usd) 
tbl.price.kwh.region.usd
```
##Price per KWh by Region and SES
###Price per KWh by Region and SES with subsidies
```{r}
#df.bill.cons$
tbl.mean.price.kwh.usd <- dstrata %>% 
  group_by(SES, region.f)  %>%
  summarize(mean_price_kwh_usd = survey_mean(price.kwh.notax.usd, na.rm= TRUE, vartype = c("ci"))) 

tbl.mean.price.kwh.usd <- na.omit(tbl.mean.price.kwh.usd) 
tbl.mean.price.kwh.usd
```

```{r}
#include estimation for all regions
tbl.price.kwh.region.usd['SES'] = 'All'
tbl.mean.price.kwh.usd <- rbind(tbl.mean.price.kwh.usd, tbl.price.kwh.region.usd)

tbl.mean.price.kwh.usd$SES <- factor(tbl.mean.price.kwh.usd$SES,
                               levels = c('Low', 'Mid-Low', 'Mid-High', 'High', 'All'),
                               labels = c('Low', 'Mid-Low', 'Mid-High', 'High', 'All'))

tbl.mean.price.kwh.usd$region.f <- factor(tbl.mean.price.kwh.usd$region.f,
                               levels = c('Extremely hot', 'Temperate', 'Tropical', 'All'),
                               labels = c('Extremely hot', 'Temperate', 'Tropical', 'All'))

tbl.mean.price.kwh.usd <- tbl.mean.price.kwh.usd %>%
  rename(Region = region.f)

#write.csv(tbl.mean.cons.ses,"./output/tables/tbl.mean.cons.ses.month.csv", row.names = TRUE)
tbl.mean.price.kwh.usd
```

#Consumption by SES
```{r}
#Make Bar Plot
plot.price.kwh <- ggplot(tbl.mean.price.kwh.usd, aes(x = Region, y = mean_price_kwh_usd, fill=SES)) +
  geom_bar(stat="identity",position="dodge") + 
  scale_colour_viridis_d(begin = 0.1, end = 0.9, option = "inferno", aesthetics = c("fill")) +
  scale_x_discrete(name= "") +
  scale_y_continuous(name= "Electricity Price [USD/kWh]", breaks=c(200, 400, 600, 800, 1000)) +
  geom_errorbar(position=position_dodge(width=.9), width=.5, color="grey50") + 
  aes(ymin=mean_price_kwh_usd_low, ymax=mean_price_kwh_usd_upp) +
  geom_text(aes(label=round(mean_price_kwh_usd, 3)), vjust=-0.2, color="black", 
            position = position_dodge(.9), size=3)

plot.price.kwh <- plot.price.kwh + theme(axis.text.x = element_text(angle = 90, hjust = 1), 
                        panel.background = element_rect(fill = "white", colour = NA),
                        panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "grey92"), plot.title = element_text(hjust = 0.5))  +  
  labs(title = "Price of Electricity per Region and SES in Mexico 2017-2018 (after subsidies)", 
       caption ="")

plot.price.kwh

#ggsave(plot = plot.price.kwh, width = 12, height = 6, dpi = 400, 
#       filename = "./output/fig/price_kwh_SES_region_after_subsidies.png")

```

#Total cost of electricity without subsidies 

These calculations don't include fees and taxes

```{r}
number_households = sum(df.bill.cons$factor_sem) 
#million of households, according to survey
print("Million of households: ")
number_households / 1000000
```


```{r}
#million of dollars per year
# No subsidy
(number_households * tbl.kwh.price.nosub.usd$mean_price_kwh_usd * mean(df.bill.cons$cons.day.kWh.med1, na.rm = TRUE)*365 )/ 1000000

#with subsidy
(number_households * tbl.kwh.price.usd$mean_price_kwh_usd * mean(df.bill.cons$cons.day.kWh.med1, na.rm = TRUE)*365 )/ 1000000
```

```{r}
tbl.kwh.price.nosub.usd$mean_price_kwh_usd
tbl.kwh.price.usd$mean_price_kwh_usd
```


```{r}
#total GWh consumed
(sum(df.bill.cons$factor_sem)* mean(df.bill.cons$cons.day.kWh.med1, na.rm = TRUE)*365 )/ 1000000
```

```{r}

df.bill.cons['bill.tax.fee1.usd'] <- df.bill.cons$bill.amount1 - df.bill.cons$bill.avg1.notax / DOLLAR_TO_MXN
df.bill.cons['bill.amount.nosubsidy.usd'] <-  df.bill.cons$real.price.kwh.usd * df.bill.cons$cons.kWh.med1
df.bill.cons['bill.amount.withsubsidy.usd'] <-  df.bill.cons$price.kwh.notax * df.bill.cons$cons.kWh.med1 / DOLLAR_TO_MXN

df.bill.cons['bill.subsidy1.usd'] <- df.bill.cons$bill.amount.nosubsidy.usd - (df.bill.cons$bill.avg1.notax / DOLLAR_TO_MXN)


df.bill.temp <- subset(df.bill.cons,
                        select=c('bill.tax.fee1.usd', 'bill.amount1', 'real.price.kwh.usd', 'cons.kWh.med1',  'price.kwh.notax',
                                 'bill.avg1.notax', 'bill.amount.nosubsidy.usd', 
                                 'bill.amount.withsubsidy.usd', 'bill.subsidy1.usd','factor_sem'))


stat.costs <- descr(df.bill.temp, 
                           weights = df.bill.temp$factor_sem, 
                           round.digits = 3,
                           stats = c("mean", "sd", "min", 
                                     "med", "max", "pct.valid"), 
                           transpose = TRUE)

stat.costs <- as.data.frame(stat.costs)

knitr::kable(stat.costs, 
      label = 'table-stats-kwh-price', digits = 3, align = 'c', booktabs = TRUE,
  caption = 'Summary statistics') %>%
  kable_styling(bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive"),
                fixed_thead = T) %>%
  scroll_box(width = "100%")
```
```{r}
#values that have a bill below the minimum charge are discarded
df.bill.cons$bill.avg1.notax[df.bill.cons$bill.avg1.notax <= min(df.tariff.scheme$cost.kwh)*25] <- NA  

df.bill.cons.sub <- subset(df.bill.cons, bill.avg1.notax	<= min(df.tariff.scheme$cost.kwh)*25)
df.bill.cons.sub
```

```{r}
#23.731 / 39.28

```
