---
title: "Analysis of ENCEVI 2018"
author: '[Mauricio Hernandez](http://mauricioh2.com)'
date: "`r Sys.Date()`"
subtitle: Estimate Subsidies
bibliography: energy_mex.bib
---


```{r, include=FALSE}
#Remove all objects from current workspace and call garbage collector
rm(list=ls())
gc()

source("./script/general_setup_functions.R")
source("./script/compute_consumption_bill.R")
source("./script/get_bill_periodicity.R")
```


Notes: Dollar Exchange Rate: 	18.70	15-Mar-18
```{r}
DOLLAR_TO_MXN <- 18.70
```

```{r knitr_init, include=FALSE}
#options(max.print = "75")
opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               comment = NA,
               message = FALSE,
               warning = FALSE,
               fig.path = 'figures/')
```

```{r setup, include=FALSE}
#Generates the css needed by summarytools
st_css()

#devtools::install_github("ropensci/plotly")
knitr::opts_chunk$set(fig.width = 8, collapse = TRUE)

knitr::knit_hooks$set(output = function(x, options){
  if(!is.null(options$max_height)){
    paste('<pre style = "max-height:',
          options$max_height, 
          ';float: left; width: 910px; overflow-y: auto;">',
          x,
          "</pre>",
          sep = "")
  } else{
    x
  }
})
```

**Loading datasets**
</br>This dataset was obtained with the "06_getting_consumption_statistics.Rmd" notebook.
```{r}

df.bill.cons <- read.csv("./output/df_bill_wconsump.csv", 
                           header = TRUE, na.strings="")

df.tariff.scheme  <- read.csv("./output/tables/tariff_cost.csv", 
                              header = TRUE, na.strings="")

df.dac.scheme  <- read.csv("./output/tables/dac_tariff_scheme.csv", 
                              header = TRUE, na.strings="")
head(df.dac.scheme)
```


```{r}
df.dac.scheme <- df.dac.scheme %>%
  rename(is.summer1 = Summer,
         final.month.mod = month,
         dac.price.kwh = cost.kwh,
         dac.fixed.charge = fixed.charge)

names(df.dac.scheme)[1] <- "tariff"

df.dac.scheme <- subset(df.dac.scheme, select=-c(tariff,is.summer1, region, final.month.mod, year))
head(df.dac.scheme)
```


```{r}
#Merging datasets
dim(df.bill.cons)
df.bill.cons.dac <- merge(df.bill.cons, df.dac.scheme, 
                          by = c("vlookupindex"), all.x = TRUE) 
dim(df.bill.cons.dac)


```

## Payments per KWh before taxes
```{r}
df.bill.cons.dac$price.kwh.notax <- df.bill.cons.dac$bill.avg1.notax / df.bill.cons$cons.kWh.med1
df.bill.cons.dac$price.kwh.notax.usd <- df.bill.cons.dac$price.kwh.notax / DOLLAR_TO_MXN
df.bill.cons.dac$dac.price.kwh.usd <-  df.bill.cons.dac$dac.price.kwh / DOLLAR_TO_MXN
df.bill.cons.dac$dac.fixed.charge.usd <- df.bill.cons.dac$dac.fixed.charge / DOLLAR_TO_MXN
```

##Price per KWh
###Price without subsidies
```{r}
options( survey.lonely.psu = "adjust" )
dstrata <- df.bill.cons.dac %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl.kwh.price.nosub.usd <- dstrata %>%
  summarize(mean_price_kwh_usd = survey_mean(dac.price.kwh.usd, na.rm= TRUE, vartype = c("se"))) 

tbl.kwh.price.nosub.usd
```

###Price with subsidies
```{r}
options(survey.lonely.psu = "adjust" )
dstrata <- df.bill.cons.dac %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl.kwh.price.usd <- dstrata %>%
  summarize(mean_price_kwh_usd = survey_mean(price.kwh.notax.usd, na.rm= TRUE, vartype = c("se"))) 

tbl.kwh.price.usd
```

##Price per KWh by SES
### Price without subsidies
```{r}
tbl.price.kwh.ses.nosub.usd <- dstrata %>% 
  group_by(SES)  %>%
  summarize(mean_price_kwh_usd = survey_mean(dac.price.kwh.usd, na.rm= TRUE, vartype = c("ci"))) 

tbl.price.kwh.ses.nosub.usd <- na.omit(tbl.price.kwh.ses.nosub.usd) 
#%>%   rename(c('socioec status'= 'stat.socio'))

tbl.price.kwh.ses.nosub.usd$SES <- as.character(tbl.price.kwh.ses.nosub.usd$SES)

tbl.price.kwh.ses.nosub.usd
```

### Price with subsidies
```{r}
tbl.price.kwh.ses.usd <- dstrata %>% 
  group_by(SES)  %>%
  summarize(mean_price_kwh_usd = survey_mean(price.kwh.notax.usd, na.rm= TRUE, vartype = c("ci"))) 

tbl.price.kwh.ses.usd <- na.omit(tbl.price.kwh.ses.usd) 
#%>%   rename(c('socioec status'= 'stat.socio'))

tbl.price.kwh.ses.usd$SES <- as.character(tbl.price.kwh.ses.usd$SES)

tbl.price.kwh.ses.usd
```

##Price per KWh by Region
### Price of KWh per Region without subsidies
```{r}
tbl.price.kwh.region.nosub.usd <- dstrata %>% 
  group_by(region.f)  %>%
  summarize(mean_price_kwh_usd = survey_mean(dac.price.kwh.usd, na.rm= TRUE, vartype = c("ci"))) 

tbl.price.kwh.region.nosub.usd <- na.omit(tbl.price.kwh.region.nosub.usd) 
#%>%   rename(c('socioec status'= 'stat.socio'))

#tbl.price.kwh.region.usd$SES <- as.character(tbl.price.kwh.region.usd$SES)

tbl.price.kwh.region.nosub.usd
```

### Price of KWh per Region with subsidies
```{r}
#df.bill.cons.dac$
tbl.price.kwh.region.usd <- dstrata %>% 
  group_by(region.f)  %>%
  summarize(mean_price_kwh_usd = survey_mean(price.kwh.notax.usd, na.rm= TRUE, vartype = c("ci"))) 

tbl.price.kwh.region.usd <- na.omit(tbl.price.kwh.region.usd) 
tbl.price.kwh.region.usd
```
##Price per KWh by Region and SES
###Price per KWh by Region and SES with subsidies
```{r}
#df.bill.cons.dac$
tbl.mean.price.kwh.usd <- dstrata %>% 
  group_by(SES, region.f)  %>%
  summarize(mean_price_kwh_usd = survey_mean(price.kwh.notax.usd, na.rm= TRUE, vartype = c("ci"))) 

tbl.mean.price.kwh.usd <- na.omit(tbl.mean.price.kwh.usd) 
tbl.mean.price.kwh.usd
```

```{r}
#include estimation for all regions
tbl.price.kwh.region.usd['SES'] = 'All'
tbl.mean.price.kwh.usd <- rbind(tbl.mean.price.kwh.usd, tbl.price.kwh.region.usd)

tbl.mean.price.kwh.usd$SES <- factor(tbl.mean.price.kwh.usd$SES,
                               levels = c('Low', 'Mid-Low', 'Mid-High', 'High', 'All'),
                               labels = c('Low', 'Mid-Low', 'Mid-High', 'High', 'All'))

tbl.mean.price.kwh.usd$region.f <- factor(tbl.mean.price.kwh.usd$region.f,
                               levels = c('Extremely hot', 'Temperate', 'Tropical', 'All'),
                               labels = c('Extremely hot', 'Temperate', 'Tropical', 'All'))

tbl.mean.price.kwh.usd <- tbl.mean.price.kwh.usd %>%
  rename(Region = region.f)

#write.csv(tbl.mean.cons.ses,"./output/tables/tbl.mean.cons.ses.month.csv", row.names = TRUE)
tbl.mean.price.kwh.usd
```

#Consumption by SES
```{r}
#Make Bar Plot
plot.price.kwh <- ggplot(tbl.mean.price.kwh.usd, aes(x = Region, y = mean_price_kwh_usd, fill=SES)) +
  geom_bar(stat="identity",position="dodge") + 
  scale_colour_viridis_d(begin = 0.1, end = 0.9, option = "inferno", aesthetics = c("fill")) +
  scale_x_discrete(name= "") +
  scale_y_continuous(name= "Electricity Price [USD/kWh]", breaks=c(200, 400, 600, 800, 1000)) +
  geom_errorbar(position=position_dodge(width=.9), width=.5, color="grey50") + 
  aes(ymin=mean_price_kwh_usd_low, ymax=mean_price_kwh_usd_upp) +
  geom_text(aes(label=round(mean_price_kwh_usd, 3)), vjust=-0.2, color="black", 
            position = position_dodge(.9), size=3)

plot.price.kwh <- plot.price.kwh + theme(axis.text.x = element_text(angle = 90, hjust = 1), 
                        panel.background = element_rect(fill = "white", colour = NA),
                        panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "grey92"), plot.title = element_text(hjust = 0.5))  +  
  labs(title = "Price of Electricity per Region and SES in Mexico 2017-2018 (after subsidies)", 
       caption ="")

plot.price.kwh

ggsave(plot = plot.price.kwh, width = 12, height = 6, dpi = 400, 
       filename = "./output/fig/price_kwh_SES_region_after_subsidies.png")

```

#Total cost of electricity without subsidies 

These calculations don't include fees and taxes

```{r}
(sum(df.bill.cons.dac$factor_sem) * 413)/(1000000)
```

```{r}

df.bill.cons.dac['bill.tax.fee1'] <- df.bill.cons.dac$bill.amount1 - df.bill.cons.dac$bill.avg1.notax
df.bill.cons.dac['bill.amount.nosubsidy'] <-  df.bill.cons.dac$dac.price.kwh * df.bill.cons$cons.kWh.med1
df.bill.cons.dac['bill.amount.withsubsidy'] <-  df.bill.cons.dac$price.kwh.notax * df.bill.cons$cons.kWh.med1

df.bill.cons.dac['bill.subsidy1'] <- df.bill.cons.dac$bill.amount.nosubsidy - df.bill.cons.dac$bill.avg1.notax

df.bill.temp <- subset(df.bill.cons.dac,
                        select=c('bill.tax.fee1', 'bill.amount1', 
                                 'dac.price.kwh.usd', 'cons.kWh.med1', 'dac.price.kwh', 'price.kwh.notax',
                                 'bill.avg1.notax', 'bill.amount.nosubsidy', 
                                 'bill.amount.withsubsidy', 'bill.subsidy1','factor_sem'))
stat.costs <- descr(df.bill.temp, 
                           weights = df.bill.temp$factor_sem, 
                           round.digits = 3,
                           stats = c("mean", "sd", "min", 
                                     "med", "max", "pct.valid"), 
                           transpose = TRUE)

stat.costs <- as.data.frame(stat.costs)

knitr::kable(stat.costs, 
      label = 'table-stats-kwh-price', digits = 3, align = 'c', booktabs = TRUE,
  caption = 'Summary statistics') %>%
  kable_styling(bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive"),
                fixed_thead = T) %>%
  scroll_box(width = "100%")
```

df.bill.cons$cons.kWh.med1
df.bill.cons.dac$cons.month.kWh.med1

df.bill.cons.dac$bill.amount1


df.bill.cons.dac$price.kwh.notax <- df.bill.cons.dac$bill.avg1.notax / df.bill.cons$cons.kWh.med1
df.bill.cons.dac$price.kwh.notax.usd <- df.bill.cons.dac$price.kwh.notax / DOLLAR_TO_MXN
df.bill.cons.dac$dac.price.kwh.usd <-  df.bill.cons.dac$dac.price.kwh / DOLLAR_TO_MXN
df.bill.cons.dac$dac.fixed.charge.usd <- df.bill.cons.dac$dac.fixed.charge / DOLLAR_TO_MXN


options(survey.lonely.psu = "adjust" )
dstrata <- df.bill.cons.dac %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 


df.bill.sub <- subset(df.bill.cons.dac,
                        select=c('cons.month.kWh.low1', 'cons.month.kWh.med1', 
                                 'cons.month.kWh.hig1', 'factor_sem'))

```



### Monthly Electricity Consumption in kWh
```{r tbl-stats-kWh-month, results="hold"}
df.bill.temp <- subset(df.enc.dwell.bill,
                        select=c('cons.month.kWh.low1', 'cons.month.kWh.med1', 
                                 'cons.month.kWh.hig1', 'factor_sem'))

stat.kWh.month <- descr(df.bill.temp, 
                           weights = df.bill.temp$factor_sem, 
                           round.digits = 3,
                           stats = c("mean", "sd", "min", 
                                     "med", "max", "pct.valid"), 
                           transpose = TRUE)

stat.kWh.month <- as.data.frame(stat.kWh.month)

knitr::kable(stat.kWh.month, 
      label = 'table-stats-kWhmonth', digits = 3, align = 'c', booktabs = TRUE,
  caption = 'Summary statistics of monthly electricity consumption (in kWh)') %>%
  kable_styling(bootstrap_options = c("striped", "hover", 
                                      "condensed", "responsive"),
                fixed_thead = T) %>%
  scroll_box(width = "100%")

rm(df.bill.temp)

stat.kWh.month$percent.valid <-stat.kWh.month$`% Total`
```

df.bill.cons$final.month.mod
