---
title: "Analysis of ENCEVI 2018"
author: '[Mauricio Hernandez](http://mauricioh2.com)'
date: "`r Sys.Date()`"
subtitle: Estimate Subsidies
bibliography: energy_mex.bib
---


```{r, include=FALSE}
#Remove all objects from current workspace and call garbage collector
rm(list=ls())
gc()

source("./script/general_setup_functions.R")
source("./script/compute_consumption_bill.R")
source("./script/get_bill_periodicity.R")
```


Notes: Dollar Exchange Rate: 	18.70	30-Jun-18, Ref. Banxico, 2021 (https://www.banxico.org.mx/tipcamb/main.do?page=tip&idioma=sp)
```{r}
DOLLAR_TO_MXN <- 19.69
```

```{r knitr_init, include=FALSE}
#options(max.print = "75")
opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               comment = NA,
               message = FALSE,
               warning = FALSE,
               fig.path = 'figures/')
```

```{r setup, include=FALSE}
#Generates the css needed by summarytools
st_css()

#devtools::install_github("ropensci/plotly")
knitr::opts_chunk$set(fig.width = 8, collapse = TRUE)

knitr::knit_hooks$set(output = function(x, options){
  if(!is.null(options$max_height)){
    paste('<pre style = "max-height:',
          options$max_height, 
          ';float: left; width: 910px; overflow-y: auto;">',
          x,
          "</pre>",
          sep = "")
  } else{
    x
  }
})
```

**Loading datasets**
</br>This dataset was obtained with the "06_getting_consumption_statistics.Rmd" notebook.
```{r}

df.bill.cons <- read.csv("./output/df_bill_wconsumpV3.csv", 
                           header = TRUE, na.strings="")

df.tariff.scheme  <- read.csv("./output/tables/tariff_cost.csv", 
                              header = TRUE, na.strings="")

head(df.tariff.scheme)
```

Actual cost of electricity is similar for each tariff and for summer and no summer seasons
The cost per Kwh is $2.802 MXN
```{r}
df.tariff.scheme.sub <- subset(df.tariff.scheme, tariff.block.name=="extra",
                               select= c(tariff, summer, cost.kwh))
df.tariff.scheme.sub
```


```{r}
# for all the tariffs the max. cost per kwh is $2.802
COST_KWH_CFE <- mean(df.tariff.scheme.sub$cost.kwh)
COST_KWH_CFE_USD <- COST_KWH_CFE / DOLLAR_TO_MXN
```

```{r}
freq(df.bill.cons$tariff)
```

```{r}
#Remove household recors whose electricity consumption couldn't be estimated
df.bill.cons.sub <- subset(df.bill.cons, (!is.na(df.bill.cons$bill_month_kwh_with_dac)))
```


##Average annual consumption
```{r}
options( survey.lonely.psu = "adjust" )
dstrata <- df.bill.cons.sub %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl.month.kwh <- dstrata %>%
  summarize(mean_month_kwh = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl.month.kwh
```

```{r}
dstrata %>% 
  summarize(mean_month_kwh = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 
```
##Number of households by tariff
```{r}
#number of households by tariff
dstrata %>% 
    group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(survey_total(grid, na.rm= TRUE, vartype = c("ci"))) 

```


##Total monthly consumption [kWh] by tariff
```{r}
tbl.total.month.kwh.tariff <- dstrata %>% 
  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(total_month_kwh = survey_total(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl.total.month.kwh.tariff <- na.omit(tbl.total.month.kwh.tariff) 

tbl.total.month.kwh.tariff
```

##Average monthly consumption [kWh] per household by tariff
```{r}
tbl_month_kwh_byhh <- dstrata %>% 
#  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(mean_month_kwh = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_kwh_byhh <- na.omit(tbl_month_kwh_byhh) 

tbl_month_kwh_byhh
```

##Average monthly cost [MXN] of electtricity by tariff

```{r}
tbl.month.cons.bytariff <- dstrata %>% 
#  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(bill_month_cost_before_tax = survey_mean(bill_month_amount_before_tax_with_dac, 
                                         na.rm= TRUE, vartype = c("ci"))) 

tbl.month.cons.bytariff <- na.omit(tbl.month.cons.bytariff) 

tbl.month.cons.bytariff
```
```{r}
tbl.month.cons.bytariff <- dstrata %>% 
#  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(bill_month_cost_before_tax = survey_mean(bill_month_amount_before_tax_with_dac, 
                                         na.rm= TRUE, vartype = c("se"))) 

tbl.month.cons.bytariff <- na.omit(tbl.month.cons.bytariff) 

tbl.month.cons.bytariff
```



##Total monthly cost [MXN] of electtricity by tariff
```{r}
tbl.total.month.cons.bytariff <- dstrata %>% 
  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(mean_month_kwh = survey_total(bill_month_amount_before_tax_with_dac, 
                                         na.rm= TRUE, vartype = c("ci"))) 

tbl.total.month.cons.bytariff <- na.omit(tbl.total.month.cons.bytariff) 

tbl.total.month.cons.bytariff
```

##Average price per kwh
```{r}
#These values don't include taxes and fees
df.bill.cons.sub$price_kwh_notax <- df.bill.cons.sub$bill_month_amount_before_tax_with_dac / df.bill.cons.sub$bill_month_kwh_with_dac
```

```{r}
options( survey.lonely.psu = "adjust" )
dstrata <- df.bill.cons.sub %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl.price.kwh <- dstrata %>%
  summarize(price_kwh = survey_mean(price_kwh_notax, na.rm= TRUE, vartype = c("ci"))) 

tbl.price.kwh
```

##Average price per kwh by Tariff
```{r}
tbl.price.kwh <- dstrata %>%
  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(price_kwh = survey_mean(price_kwh_notax, na.rm= TRUE, vartype = c("ci"))) 

tbl.price.kwh
```

## Subsidies

```{r}
df.bill.cons.sub$real.cost_bill_month <- COST_KWH_CFE * df.bill.cons.sub$bill_month_kwh_with_dac


df.bill.cons.sub$bill_subsidy_month <- df.bill.cons.sub$real.cost_bill_month - df.bill.cons.sub$bill_month_amount_before_tax_with_dac

dstrata <- df.bill.cons.sub %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl.month_subsidy <- dstrata %>%
  summarize(month_subsidy = survey_total(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 

tbl.month_subsidy
```

```{r}
  tbl_month_subsidy_byTariff <- dstrata %>%
  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(month_subsidy = survey_total(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subsidy_byTariff
```


### Monthly Subsidies per household
```{r}
tbl_month_subsidy_byhh <- dstrata %>%
  summarize(month_subsidy = survey_mean(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subsidy_byhh
```

### Monthly Subsidies per household by Tariff
```{r}
tbl_month_subsidy_byhh <- dstrata %>%
    group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(month_subsidy = survey_mean(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subsidy_byhh
```