---
title: "ENIGH 2018 - Subsidies Estimation"
author: '[Mauricio Hernandez](http://mauricioh2.com)'
date: "`r Sys.Date()`"
subtitle: Estimate Subsidies
bibliography: energy_mex.bib
---


```{r}
#Remove all objects from current workspace and call garbage collector
rm(list=ls())
gc()

source("./script/general_setup_functions.R")
#source("./script/compute_consumption_bill.R")
#source("./script/get_bill_periodicity.R")
```


```{r knitr_init, include=FALSE}
#options(max.print = "75")
opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               comment = NA,
               message = FALSE,
               warning = FALSE,
               fig.path = 'figures/')
```

```{r setup, include=FALSE}
#Generates the css needed by summarytools
st_css()

#devtools::install_github("ropensci/plotly")
knitr::opts_chunk$set(fig.width = 8, collapse = TRUE)

knitr::knit_hooks$set(output = function(x, options){
  if(!is.null(options$max_height)){
    paste('<pre style = "max-height:',
          options$max_height, 
          ';float: left; width: 910px; overflow-y: auto;">',
          x,
          "</pre>",
          sep = "")
  } else{
    x
  }
})
```


```{r}
df_enigh_hh <- read.csv("./output/enigh_hh_w_elect_v1.csv", 
                           header = TRUE, na.strings="")
df_enigh_hh <- df_enigh_hh %>%
  rename(folioviv = ï..folioviv)

head(df_enigh_hh)
```



```{r}
df.tariff.scheme  <- read.csv("./output/tables/tariff_cost.csv", 
                              header = TRUE, na.strings="")

df.tariff.scheme.sub <- subset(df.tariff.scheme, tariff.block.name=="extra",
                               select= c(tariff, summer, cost.kwh))
df.tariff.scheme.sub
```

Notes: Dollar Exchange Rate: 	18.70	30-Jun-18, Ref. Banxico, 2021 (https://www.banxico.org.mx/tipcamb/main.do?page=tip&idioma=sp)
```{r}
DOLLAR_TO_MXN <- 18.70

# for all the tariffs the max. cost per kwh is $2.802
COST_KWH_CFE <- mean(df.tariff.scheme.sub$cost.kwh)
COST_KWH_CFE_USD <- COST_KWH_CFE / DOLLAR_TO_MXN
```

```{r}
freq(df_enigh_hh$tariff_merge_manual_dac_correct)
```


```{r}
#Remove household records whose electricity consumption couldn't be estimated
df_enigh_hh_sub <- subset(df_enigh_hh, (!is.na(df_enigh_hh$bill_month_kwh_with_dac)))
```


```{r}

options(survey.lonely.psu = "adjust")
dstrata <- df_enigh_hh_sub %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor) 
```


```{r}
tbl.month.cons.bill <- dstrata %>%
  summarize(mean_month_spend = survey_mean(bill_month_amount_before_tax_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl.month.cons.bill
```

##Number of households by tariff
```{r}
#number of households by tariff
dstrata %>% 
    group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(survey_total(Nhog, na.rm= TRUE, vartype = c("ci"))) 
```

##Total monthly consumption [kWh]
```{r}
tbl.total.month.kwh <- dstrata %>% 
  summarize(total_month_kwh = survey_total(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl.total.month.kwh <- na.omit(tbl.total.month.kwh) 

tbl.total.month.kwh
```

###Total monthly consumption [kWh] by tariff
```{r}
tbl.total.month.kwh.tariff <- dstrata %>% 
  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(total_month_kwh = survey_total(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl.total.month.kwh.tariff <- na.omit(tbl.total.month.kwh.tariff) 

tbl.total.month.kwh.tariff
```

##Average monthly consumption
```{r}
tbl.month.kwh <- dstrata %>%
  summarize(mean_month_kwh = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl.month.kwh
```

##Average monthly consumption [kWh] per household by tariff
```{r}
tbl_month_kwh_byhh <- dstrata %>% 
  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(mean_month_kwh = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_kwh_byhh <- na.omit(tbl_month_kwh_byhh) 

tbl_month_kwh_byhh
```

##Average monthly cost [MXN] of electtricity by tariff
```{r}
tbl.month.cons.bytariff <- dstrata %>% 
  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(bill_month_cost_before_tax = survey_mean(bill_month_amount_before_tax_with_dac, 
                                         na.rm= TRUE, vartype = c("ci"))) 

tbl.month.cons.bytariff <- na.omit(tbl.month.cons.bytariff) 

tbl.month.cons.bytariff
```

##Total monthly cost [MXN] of electtricity by tariff
```{r}
tbl.total.month.cons.bytariff <- dstrata %>% 
  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(mean_month_kwh = survey_total(bill_month_amount_before_tax_with_dac, 
                                         na.rm= TRUE, vartype = c("ci"))) 

tbl.total.month.cons.bytariff <- na.omit(tbl.total.month.cons.bytariff) 

tbl.total.month.cons.bytariff
```

##Average price per kwh
```{r}
#These values don't include taxes and fees
df_enigh_hh_sub$price_kwh_notax <- df_enigh_hh_sub$bill_month_amount_before_tax_with_dac / df_enigh_hh_sub$bill_month_kwh_with_dac

df_enigh_hh_sub$real.cost_bill_month <- COST_KWH_CFE * df_enigh_hh_sub$bill_month_kwh_with_dac

df_enigh_hh_sub$bill_subsidy_month <- df_enigh_hh_sub$real.cost_bill_month - df_enigh_hh_sub$bill_month_amount_before_tax_with_dac


options( survey.lonely.psu = "adjust" )
dstrata <- df_enigh_hh_sub %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor) 

tbl.price.kwh <- dstrata %>%
  summarize(price_kwh = survey_mean(price_kwh_notax, na.rm= TRUE, vartype = c("ci"))) 

tbl.price.kwh
```

##Average price per kwh by Tariff
```{r}
tbl.price.kwh <- dstrata %>%
  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(price_kwh = survey_mean(price_kwh_notax, na.rm= TRUE, vartype = c("ci"))) 

tbl.price.kwh
```

## Subsidies

```{r}
tbl.month_subsidy <- dstrata %>%
  summarize(month_subsidy = survey_total(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 

tbl.month_subsidy
```

```{r}
tbl_month_subsidy_byTariff <- dstrata %>%
  group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(month_subsidy = survey_total(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subsidy_byTariff
```


### Monthly Subsidies per household
```{r}
tbl_month_subsidy_byhh <- dstrata %>%
  summarize(month_subsidy = survey_mean(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subsidy_byhh
```

### Monthly Subsidies per household by Tariff
```{r}
tbl_month_subsidy_byhh <- dstrata %>%
    group_by(tariff_merge_manual_dac_correct)  %>%
  summarize(month_subsidy = survey_mean(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subsidy_byhh
```

# Statistics by Income Decile

## Electricity Consumption by Income Decile
```{r}
tbl_month_kwh_bydecile <- dstrata %>%
  group_by(income_decile)  %>%
  summarize(kwh_decile = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_kwh_bydecile
```

## Electricity expenses by decile
```{r}
tbl_month_consump_bydecile <- dstrata %>%
  group_by(income_decile)  %>%
  summarize(cons_amount_decile = survey_mean(elect_month_spend, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_consump_bydecile
```

## Subsidies by Income Decile
```{r}
df_enigh_hh_sub$perc_spend_before_tax <- df_enigh_hh_sub$bill_month_amount_before_tax_with_dac / (df_enigh_hh_sub$ing_cor/3)
df_enigh_hh_sub$perc_spend <- df_enigh_hh_sub$elect_month_spend / (df_enigh_hh_sub$ing_cor/3)


df_enigh_hh_sub$perc_spend_no_subsidy <- (df_enigh_hh_sub$elect_month_spend +  df_enigh_hh_sub$bill_subsidy_month )/ (df_enigh_hh_sub$ing_cor/3) 

df_enigh_hh_sub2 <- subset(df_enigh_hh_sub, ing_cor!=0)
df_enigh_hh_sub2 <- subset(df_enigh_hh_sub2, !is.na(df_enigh_hh_sub2$bill_subsidy_month))

options(survey.lonely.psu = "adjust")
dstrata <- df_enigh_hh_sub2 %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor) 
```



```{r}

mean(df_enigh_hh_sub2$perc_spend)
median(df_enigh_hh_sub2$perc_spend)
mean(df_enigh_hh_sub2$perc_spend_no_subsidy)
median(df_enigh_hh_sub2$perc_spend_no_subsidy)
```





## Monthly Subsidies by Decile
```{r}
tbl_month_subs_bydecil <- dstrata %>%
  group_by(income_decile)  %>%
  summarize(subsidy = survey_mean(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subs_bydecil
```

###Percentage of electricity expenditure relative to income (with subsidies)
```{r}
tbl_month_perc_spend_bydecil <- dstrata %>%
  group_by(income_decile)  %>%
  summarize(perc_spend = survey_mean(perc_spend, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_perc_spend_bydecil
```




```{r}
tbl_month_perc_spend_region_bydecil <- dstrata %>%
  group_by(income_decile, region_f)  %>%
  summarize(perc_spend = survey_mean(perc_spend, na.rm= TRUE, vartype = c("ci"))) 
tbl_month_perc_spend_region_bydecil
```
```{r}
write.csv(tbl_month_perc_spend_region_bydecil,"./output/tables/enigh_month_perc_spend_region_bydecil.csv", row.names = FALSE)
```

## Median Values
```{r}
dstrata %>%
  summarize(perc_spend = survey_median(perc_spend, na.rm= TRUE, vartype = c("ci"))) 
```

```{r}
tbl_month_perc_spend_region_bydecil <- dstrata %>%
  group_by(income_decile, region_f)  %>%
  summarize(perc_spend = survey_median(perc_spend, na.rm= TRUE, vartype = c("ci"))) 
tbl_month_perc_spend_region_bydecil
```

```{r}
write.csv(tbl_month_perc_spend_region_bydecil,"./output/tables/enigh_median_perc_spend_region_bydecil.csv", row.names = FALSE)
```

###Percentage of electricity subsidy relative to income by region (without subsidies)
```{r}
tbl_month_perc_spend_nosubsidy_region_bydecil <- dstrata %>%
  group_by(income_decile, region_f)  %>%
  summarize(perc_spend_no_subsidy = survey_mean(perc_spend_no_subsidy, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_perc_spend_nosubsidy_region_bydecil
```
```{r}
write.csv(tbl_month_perc_spend_nosubsidy_region_bydecil,"./output/tables/enigh_month_perc_spend_nosubsidy_region_bydecil.csv", row.names = FALSE)
```

## Median values

```{r}
tbl_month_perc_spend_nosubsidy_region_bydecil <- dstrata %>%
  group_by(income_decile, region_f)  %>%
  summarize(perc_spend_no_subsidy = survey_median(perc_spend_no_subsidy, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_perc_spend_nosubsidy_region_bydecil
```

```{r}
 dstrata %>%
  summarize(perc_spend_no_subsidy = survey_median(perc_spend_no_subsidy, na.rm= TRUE, vartype = c("ci"))) 
```

```{r}
write.csv(tbl_month_perc_spend_nosubsidy_region_bydecil,"./output/tables/enigh_median_perc_spend_nosubsidy_region_bydecil.csv", row.names = FALSE)
```




df_enigh_hh_sub$perc_spend_no_subsidy <- (df_enigh_hh_sub$elect_month_spend +  df_enigh_hh_sub$bill_subsidy_month )/ (df_enigh_hh_sub$ing_cor/3) 

df_enigh_hh_sub2 <- subset(df_enigh_hh_sub, ing_cor!=0)
df_enigh_hh_sub2 <- subset(df_enigh_hh_sub2, !is.na(df_enigh_hh_sub2$bill_subsidy_month))

###Percentage of electricity subsidy relative to income by region 
```{r}
tbl_month_perc_spend_region_bydecil <- dstrata %>%
  group_by(income_decile, region_f)  %>%
  summarize(perc_spend = survey_mean(perc_spend, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_perc_spend_region_bydecil
```



```{r}
tbl_month_perc_spend_nosubsidy_region_bydecil
```
```{r}
plot.cons.month <- ggplot(tbl_month_perc_spend_region_bydecil, aes(x = income_decile, y = perc_spend*100, fill= region_f)) +
    geom_bar(stat="identity",position="dodge") + 
  scale_colour_viridis_d(begin = 0.8, end = 0.2,   option = "magma", aesthetics = c("fill")) +
    scale_x_discrete(name= c("1st              \t    2nd            \t      3rd             \t       4th             \t      5th           \t       6th           \t      7th           \t         8th               \t     9th             \t     10th")) +
    scale_y_continuous(name= "Percentage", breaks=c(0, 2, 4, 6, 8, 10), limits= c(0, 11.5)) +
  geom_errorbar(position=position_dodge(width=.9), width=.5, color="grey50") + 
  aes(ymin=perc_spend_low*100, ymax=perc_spend_upp*100) +
  geom_text(aes(label=round(perc_spend*100, 2)), vjust=-0.6, color="black", 
            position = position_dodge(.9), size=4.5) +
    facet_wrap(~ region_f, ncol = 1, scales = "free_y");

plot.cons.month <- plot.cons.month + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size=16), 
        axis.text.y = element_text(angle = 90, size=14),
        axis.title.y = element_text(angle = 90, size=16),
        plot.title = element_text(hjust = 0.5, size =18), 
        panel.background = element_rect(fill = "white", colour = NA),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "grey92"), 
        legend.position = "none", strip.text = element_text(size = 14))  +  
  labs(title = "Average Percent of Income Spent on Electricity by Income Decile and Climate Region in Mexico",
       subtitle ="Summer Season, 2018")

plot.cons.month

ggsave(plot = plot.cons.month, width = 12, height = 8, dpi = 400, 
       filename = "./output/fig/month_perc_spend_region_bydecil.png")
```



```{r}
plot.spend.month <- ggplot(tbl_month_perc_spend_nosubsidy_region_bydecil, 
                          aes(x = income_decile, y = perc_spend_no_subsidy*100, fill= region_f)) +
    geom_bar(stat="identity",position="dodge") + 
  scale_colour_viridis_d(begin = 0.8, end = 0.2,   option = "magma", aesthetics = c("fill")) +
    scale_x_discrete(name= c("1st              \t    2nd            \t      3rd             \t       4th             \t      5th           \t       6th           \t      7th           \t         8th               \t     9th             \t     10th")) +
    scale_y_continuous(name= "Percentage", breaks=c(0, 5, 10, 15, 20, 25), limits= c(0, 26.0)) +
  geom_errorbar(position=position_dodge(width=.9), width=.5, color="grey50") + 
  aes(ymin=perc_spend_no_subsidy_low*100, ymax=perc_spend_no_subsidy_upp*100) +
  geom_text(aes(label=round(perc_spend_no_subsidy*100, 2)), vjust=-0.6, color="black", 
            position = position_dodge(.9), size=4.5) +
    facet_wrap(~ region_f, ncol = 1, scales = "free_y");

plot.spend.month <- plot.spend.month + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size=16), 
        axis.text.y = element_text(angle = 90, size=14),
        axis.title.y = element_text(angle = 90, size=16),
        plot.title = element_text(hjust = 0.5, size =18), 
        panel.background = element_rect(fill = "white", colour = NA),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "grey92"), 
        legend.position = "none", strip.text = element_text(size = 14))  +  
  labs(title = "Average Percent of Income Spent on Electricity by Income Decile and Climate Region without Subsidies",
       subtitle ="Summer Season, 2018")

plot.spend.month

ggsave(plot = plot.spend.month, width = 12, height = 8, dpi = 400, 
       filename = "./output/fig/month_perc_spend_no_subsidies_region_bydecil.png")
```


```{r}
#this table was created manually
tbl_month_month_perc_spend_region_both_bydecil <- read.csv("./output/tables/enigh_month_perc_spend_region_both_bydecil.csv", 
                           header = TRUE, na.strings="")

```


```{r}
plot.cons.month <- ggplot(tbl_month_month_perc_spend_region_both_bydecil, aes(x = income_decile, y = perc_spend*100, fill= subsidized)) +
    geom_bar(stat="identity",position="dodge") + 
  scale_colour_viridis_d(begin = 0.8, end = 0.2,   option = "magma", aesthetics = c("fill")) +
    scale_x_discrete(name= c("1st              \t    2nd            \t      3rd             \t       4th             \t      5th           \t       6th           \t      7th           \t         8th               \t     9th             \t     10th")) +
    scale_y_continuous(name= "Percentage", breaks=c(0, 5, 10, 15, 20, 25), limits= c(0, 25.5)) +
  geom_errorbar(position=position_dodge(width=.9), width=.5, color="grey50") + 
  aes(ymin=perc_spend_low*100, ymax=perc_spend_upp*100) +
  geom_text(aes(label=round(perc_spend*100, 2)), vjust=-0.6, color="black", 
            position = position_dodge(.9), size=4.5) +
    facet_wrap(~ region_f, ncol = 1, scales = "free_y");

plot.cons.month <- plot.cons.month + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size=16), 
        axis.text.y = element_text(angle = 90, size=14),
        axis.title.y = element_text(angle = 90, size=16),
        plot.title = element_text(hjust = 0.5, size =18), 
        panel.background = element_rect(fill = "white", colour = NA),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "grey92"), 
        legend.position = "none", strip.text = element_text(size = 14))  +  
  labs(title = "Average Percent of Income Spent on Electricity by Income Decile and Climate Region in Mexico",
       subtitle ="Summer Season, 2018")

plot.cons.month

ggsave(plot = plot.cons.month, width = 12, height = 8, dpi = 400, 
       filename = "./output/fig/month_perc_spend_subsidy_both_region_bydecil.png")
```

# Plots with Median values


```{r}
dstrata %>%
  summarize(perc_spend = survey_median(perc_spend, na.rm= TRUE, vartype = c("ci"))) 
```


```{r}
#this table was created manually
tbl_month_month_perc_spend_region_both_bydecil <- read.csv("./output/tables/enigh_median_perc_spend_region_both_bydecil.csv", 
                           header = TRUE, na.strings="")

```


```{r}
plot.cons.month <- ggplot(tbl_month_month_perc_spend_region_both_bydecil, aes(x = income_decile, y = perc_spend*100, fill= subsidized)) +
    geom_bar(stat="identity",position="dodge") + 
  scale_colour_viridis_d(begin = 0.8, end = 0.2,   option = "magma", aesthetics = c("fill")) +
    scale_x_discrete(name= c("1st              \t    2nd            \t      3rd             \t       4th             \t      5th           \t       6th           \t      7th           \t         8th               \t     9th             \t     10th")) +
    scale_y_continuous(name= "Percentage", breaks=c(0, 5, 10, 15, 20), limits= c(0, 20.5)) +
  geom_errorbar(position=position_dodge(width=.9), width=.5, color="grey50") + 
  aes(ymin=perc_spend_low*100, ymax=perc_spend_upp*100) +
  geom_text(aes(label=round(perc_spend*100, 2)), vjust=-0.6, color="black", 
            position = position_dodge(.9), size=4.5) +
    facet_wrap(~ region_f, ncol = 1, scales = "free_y");

plot.cons.month <- plot.cons.month + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size=16), 
        axis.text.y = element_text(angle = 90, size=14),
        axis.title.y = element_text(angle = 90, size=16),
        plot.title = element_text(hjust = 0.5, size =18), 
        panel.background = element_rect(fill = "white", colour = NA),
        panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "grey92"), 
        legend.position = "none", strip.text = element_text(size = 14))  +  
  labs(title = "Median Percent of Income Spent on Electricity by Income Decile and Climate Region in Mexico",
       subtitle ="Summer Season, 2018")

plot.cons.month

ggsave(plot = plot.cons.month, width = 12, height = 8, dpi = 400, 
       filename = "./output/fig/median_perc_spend_subsidy_both_region_bydecil.png")
```


#Estimations by Climate Region

*Electricity consumption by income decile and climate region*
```{r}

tbl_month_kwh_bydec_region <- dstrata %>%
  group_by(income_decile, region_f)  %>%
  summarize(kwh_decile = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_kwh_bydec_region
```
```{r}
write.csv(tbl_month_kwh_bydec_region,"./output/tables/enigh_cons_by_income_decil_region.csv", row.names = FALSE)
```

#Estimations by State

Electricity consumption by state

```{r}
tbl_month_kwh_bystate <- dstrata %>%
  group_by(state.f, is_summer1)  %>%
  summarize(kwh_month = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_kwh_bystate
```

```{r}
write.csv(tbl_month_kwh_bystate,"./output/tables/enigh_mean_cons_state.csv", row.names = FALSE)
```