---
title: "ENCEVI 2018 - Subsidies By SES"
author: '[Mauricio Hernandez](http://mauricioh2.com)'
date: "`r Sys.Date()`"
subtitle: Estimate Subsidies
bibliography: energy_mex.bib
---


```{r}
#Remove all objects from current workspace and call garbage collector
rm(list=ls())
gc()

source("./script/general_setup_functions.R")
#source("./script/compute_consumption_bill.R")
#source("./script/get_bill_periodicity.R")
```


```{r knitr_init, include=FALSE}
#options(max.print = "75")
opts_chunk$set(echo = TRUE,
	             cache = FALSE,
               prompt = FALSE,
               tidy = FALSE,
               comment = NA,
               message = FALSE,
               warning = FALSE,
               fig.path = 'figures/')
```

```{r setup, include=FALSE}
#Generates the css needed by summarytools
st_css()

#devtools::install_github("ropensci/plotly")
knitr::opts_chunk$set(fig.width = 8, collapse = TRUE)

knitr::knit_hooks$set(output = function(x, options){
  if(!is.null(options$max_height)){
    paste('<pre style = "max-height:',
          options$max_height, 
          ';float: left; width: 910px; overflow-y: auto;">',
          x,
          "</pre>",
          sep = "")
  } else{
    x
  }
})
```


```{r}
df_encevi_hh <- read.csv("./output/df_bill_wconsump.csv", 
                           header = TRUE, na.strings="")
df_encevi_hh <- df_encevi_hh %>%
  rename(folioviv = ï..folio)

head(df_encevi_hh)
```


```{r}
df.tariff.scheme  <- read.csv("./output/tables/tariff_cost.csv", 
                              header = TRUE, na.strings="")

df.tariff.scheme.sub <- subset(df.tariff.scheme, tariff.block.name=="extra",
                               select= c(tariff, summer, cost.kwh))
df.tariff.scheme.sub
```

Notes: Dollar Exchange Rate: 	18.70	30-Jun-18, Ref. Banxico, 2021 (https://www.banxico.org.mx/tipcamb/main.do?page=tip&idioma=sp)
```{r}
DOLLAR_TO_MXN <- 19.69

# for all the tariffs the max. cost per kwh is $2.802
COST_KWH_CFE <- mean(df.tariff.scheme.sub$cost.kwh)
COST_KWH_CFE_USD <- COST_KWH_CFE / DOLLAR_TO_MXN
```


# Average electricity consumption (in Kwh) by household SESs in 2018

```{r}
#df_encevi_hh$bill_month_kwh_with_dac
df_encevi_hh_sub <- subset(df_encevi_hh, (!is.na(df_encevi_hh$bill_month_kwh_with_dac)))
```


# Statistics by Consumption SES

## Electricity Consumption by SES
```{r}
options( survey.lonely.psu = "adjust")
dstrata <- df_encevi_hh_sub %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl_month_kwh_bySES <- dstrata %>%
  group_by(ses)  %>%
  summarize(kwh_ses = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_kwh_bySES
```


## Electricity expenses by SES
```{r}
#Amount paid after taxes
tbl_month_consump_bySES <- dstrata %>%
  group_by(ses)  %>%
  summarize(cons_amount_SES = survey_mean(bill_amount1, na.rm= TRUE, vartype = c("ci"))) 
tbl_month_consump_bySES
```


```{r}
sum(df_encevi_hh$factor_sem)
sum(df_encevi_hh_sub$factor_sem)

sum(df_encevi_hh_sub$factor_sem)

sum(df_encevi_hh$Nhog)
sum(df_encevi_hh_sub$Nhog)

sum(df_encevi_hh_sub$Nhog)/ sum(df_encevi_hh$Nhog)
sum(df_encevi_hh_sub$factor_sem) / sum(df_encevi_hh$factor_sem)
sum(df_encevi_hh$Nhog) - sum(df_encevi_hh_sub$Nhog)

freq(df_encevi_hh$tariff_merge_manual_dac_correct)
```


## Subsidies

```{r}
df_encevi_hh_sub$real.cost_bill_month <- COST_KWH_CFE * df_encevi_hh_sub$bill_month_kwh_with_dac

df_encevi_hh_sub$bill_subsidy_month <- df_encevi_hh_sub$real.cost_bill_month - df_encevi_hh_sub$bill_month_amount_before_tax_with_dac
# as some values would be negative, because of DAC tariffs
df_encevi_hh_sub$bill_subsidy_month_pos <- df_encevi_hh_sub$bill_subsidy_month
df_encevi_hh_sub$bill_subsidy_month_pos[df_encevi_hh_sub$bill_subsidy_month_pos < 0] <- 0

df_encevi_hh_sub$subsidy_per_kwh_pos <- df_encevi_hh_sub$bill_subsidy_month_pos / df_encevi_hh_sub$bill_month_kwh_with_dac

df_encevi_hh_sub$subsidy_per_kwh <- df_encevi_hh_sub$bill_subsidy_month / df_encevi_hh_sub$bill_month_kwh_with_dac

dstrata <- df_encevi_hh_sub %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl.month_subsidy <- dstrata %>%
  summarize(month_subsidy = survey_total(bill_subsidy_month_pos, na.rm= TRUE, vartype = c("ci"))) 

tbl.month_subsidy
```


## Average Monthly Subsidies
```{r}
tbl_month_subs_bydecil <- dstrata %>%
  #group_by(income_SES)  %>%
  summarize(subsidy = survey_mean(bill_subsidy_month_pos, na.rm= TRUE, vartype = c("ci"))) 
#bill_subsidy_month_pos
tbl_month_subs_bydecil
```

## Average Monthly Subsidies by Consumption SES
```{r}
tbl_month_subs_bydecil <- dstrata %>%
  group_by(ses)  %>%
  summarize(subsidy = survey_mean(bill_subsidy_month_pos, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subs_bydecil
```

## Total Monthly Subsidies
```{r}
tbl_month_subs_bydecil <- dstrata %>%
  summarize(subsidy = survey_total(bill_subsidy_month_pos, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subs_bydecil
```

## Total Monthly Subsidies by Consumption SES
```{r}
#income_SES  ses
tbl_month_subs_bydecil <- dstrata %>%
  group_by(ses)  %>%
  summarize(subsidy = survey_total(bill_subsidy_month_pos, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subs_bydecil
```



## Subsidies per KWh
```{r}


options( survey.lonely.psu = "adjust")

dstrata <- df_encevi_hh_sub %>%
  as_survey_design(id = upm, strata = est_dis, weights = factor_sem) 

tbl_month_subs_kwh <- dstrata %>%
  summarize(subsidy = survey_mean(subsidy_per_kwh_pos, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subs_kwh
```
## Subsidies per KWh by Consumption SES
```{r}
tbl_month_subs_kwh_bydecil <- dstrata %>%
  group_by(ses)  %>%
  summarize(subsidy = survey_mean(subsidy_per_kwh_pos, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subs_kwh_bydecil
```


#Using also subsidies with negative values for DAC

## Average Monthly Subsidies
```{r}
tbl_month_subs_bydecil <- dstrata %>%
  #group_by(income_SES)  %>%
  summarize(subsidy = survey_mean(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 
#bill_subsidy_month

tbl_month_subs_bydecil
```

## Average Monthly Subsidies by Consumption SES
```{r}
tbl_month_subs_bydecil <- dstrata %>%
  group_by(ses)  %>%
  summarize(subsidy = survey_mean(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subs_bydecil
```

## Total Monthly Subsidies
```{r}
tbl_month_subs_bydecil <- dstrata %>%
  summarize(subsidy = survey_total(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subs_bydecil
```

## Total Monthly Subsidies by Consumption SES
```{r}
#income_SES  ses
tbl_month_subs_bydecil <- dstrata %>%
  group_by(ses)  %>%
  summarize(subsidy = survey_total(bill_subsidy_month, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subs_bydecil
```

## Subsidies per KWh
```{r}
tbl_month_subs_kwh <- dstrata %>%
  summarize(subsidy = survey_mean(subsidy_per_kwh, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subs_kwh
```
## Subsidies per KWh by Consumption SES
```{r}
tbl_month_subs_kwh_bydecil <- dstrata %>%
  group_by(ses)  %>%
  summarize(subsidy = survey_mean(subsidy_per_kwh, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_subs_kwh_bydecil
```

```{r}
tbl_month_kwh_bySES_region_summer <- dstrata %>%
  group_by(ses, region_f, is_summer1)  %>%
  summarize(kwh_month = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_kwh_bySES_region_summer
```


#Estimations by Summer Season
```{r}
tbl_month_kwh_bySES_summer <- dstrata %>%
  group_by(ses, is_summer1)  %>%
  summarize(kwh_month = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_kwh_bySES_summer
```


*Electricity consumption by SES and climate region*
```{r}
tbl_month_kwh_bySES_region <- dstrata %>%
  group_by(ses, region_f)  %>%
  summarize(kwh_month = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_kwh_bySES_region
```

```{r}
write.csv(tbl_month_kwh_bySES_region,"./output/tables/encevi_cons_by_SES_region.csv", row.names = FALSE)
```


```{r}
tbl_month_kwh_bySES_region_summer <- dstrata %>%
  group_by(ses, region_f, is_summer1)  %>%
  summarize(kwh_month = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_kwh_bySES_region_summer
```
```{r}
write.csv(tbl_month_kwh_bySES_region_summer,"./output/tables/encevi_cons_by_SES_region_summer.csv", row.names = FALSE)
```


#Estimations by State

*Electricity consumption by state*

```{r}

#df_encevi_hh_sub$cons

tbl_month_kwh_bystate <- dstrata %>%
  group_by(state_f, is_summer1)  %>%
  summarize(kwh_month = survey_mean(bill_month_kwh_with_dac, na.rm= TRUE, vartype = c("ci"))) 

tbl_month_kwh_bystate
```

